Classification structurée pour l'apprentissage par renforcement

### 2012-09-14 klein2012structured_RIA/scirl_ria.tex
Traduction de la première sous-section

\begin{document}
\newtheorem{prop}[theorem]{Proposition}
\cite{Ng:2000}
\maketitle

@@

\section{Background and Notations}
\label{sec:background}

\subsection{(Inverse) Reinforcement Learning}
\label{subsec:background:irl}

A Markov Decision process (MDP)~\cite{Puterman:1994} is a tuple
$\{\s,\A,\p,\R,\gamma\}$ where $\s$ is the finite state
space\footnote{This work can be extended to compact state spaces, up
to some technical aspects.}, $\A$ the finite actions space, $\p =
\{P_a = (p(s'|s,a))_{1\leq s,s'\leq |\s|}, a\in\A\}$ the set of
Markovian transition probabilities, $\R\in\mathbb{R}^\s$ the
state-dependent reward function and $\gamma$ the discount factor. A
deterministic policy $\pi\in\s^\A$ defines the behavior of an agent.
The quality of this control is quantified by the value function
$v_\R^\pi\in\mathbb{R}^\s$, associating to each state the cumulative
discounted reward for starting in this state and following the
policy $\pi$ afterwards: $v_\R^\pi(s) = \E[\sum_{t\geq 0} \gamma^t
\R(S_t)|S_0=s,\pi]$. An optimal policy $\pi_\R^*$ (according to the
reward function $\R$) is a policy of associated value function
$v^*_\R$ satisfying $v_\R^* \geq v_\R^\pi$, for any policy $\pi$ and
componentwise.

Let $P_\pi$ be the stochastic matrix $P_\pi =
(p(s'|s,\pi(s)))_{1\leq s,s'\leq |\s|}$. With a slight abuse of
notation, we may write $a$ the policy which associates the action
$a$ to each state $s$. The Bellman evaluation (resp. optimality)
operators $T^\pi_\R\text{ (resp. $T^*_\R$)}:\mathbb{R}^\s
\rightarrow \mathbb{R}^\s$ are defined as $T^\pi_\R v = \R + \gamma
P_\pi v$ and $T_\R^*v = \max_\pi T_\R^\pi v$.
%\begin{equation}
%  T^\pi_\R v = \R + \gamma P_\pi v \text{ and } T_\R^*v = \max_\pi T_\R^\pi v.
%\end{equation}
These operators are contractions and $v_\R^\pi$ and $v^*_\R$ are
their respective fixed-points: $v_\R^\pi = T^\pi_\R v_\R^\pi$ and
$v^*_\R = T^*_\R v^*_\R$. The action-value function
$Q^\pi\in\mathbb{R}^{\s\times \A}$ adds a degree of freedom on the
choice of the first action, it is formally defined as $Q_\R^\pi(s,a)
= [T^a_\R v^\pi_\R](s)$. We also write $\rho_\pi$ the stationary
distribution of the policy $\pi$ (satisfying $\rho_\pi^\top P_\pi =
\rho_\pi^\top$).

Reinforcement learning and approximate dynamic programming aim at
estimating the optimal control policy $\pi^*_\R$ when the model
(transition probabilities and the reward function) is unknown (but
observed through interactions with the system to be controlled) and
when the state space is too large to allow exact representations of
the objects of interest (as value functions or
policies)~\cite{Bertsekas:1996,Sutton:1998,szepesvari2010c}.
%
We refer to this as the direct problem.
%
On the contrary, (approximate) inverse reinforcement
learning~\cite{Ng:2000} aim at estimating a reward function for
which an observed policy is (nearly) optimal.
%
Let us call this policy the expert policy, denoted $\pi_E$. We may
assume that it optimizes some unknown reward function $\R_E$. The
aim of IRL is to compute some reward $\hat{\R}$ such that the expert
policy is (close to be) optimal, that is such that $v^*_{\hat{\R}}
\approx v^{\pi_E}_{\hat{\R}}$.
%
We refer to this as the inverse problem.

Similarly to the direct problem, the state space may be too large
for the reward function to admit a practical exact representation.
Therefore, we restrict our search of a good reward among linearly
parameterized functions. Let $\phi(s) = (\phi_1(s)  \dots
\phi_p(s))^\top$
%\begin{pmatrix}
%  \phi_1(s) & \dots & \phi_p(s)
%\end{pmatrix}^\top$
be a feature vector composed of $p$ basis function
$\phi_i\in\mathbb{R}^\s$, we define the  parameterized reward
functions as $\R_\theta(s) = \theta^\top \phi(s) = \sum_{i=1}^p
\theta_i \phi_i(s)$.
%\begin{equation}
%  \R_\theta(s) = \theta^\top \phi(s) = \sum_{i=1}^p \theta_i \phi_i(s).
%\end{equation}
Searching a good reward thus reduces to searching a good parameter
vector $\theta \in\mathbb{R}^p$. Notice that we will use
interchangeably $\R_\theta$ and $\theta$ as subscripts
(\textit{e.g.}, $v_\theta^\pi$ for $v_{\R_\theta}^\pi$).
Parameterizing the reward this way implies a related
parameterization for the action-value function:
\begin{equation}
  Q^\pi_\theta(s,a) = \theta^\top \mu^\pi(s,a) \text{ with }
  \mu^\pi(s,a) = \E[\sum_{t\geq 0} \gamma^t
  \phi(S_t)|S_0=s,A_0=a,\pi].
  \label{eq:def:mu}
\end{equation}
Therefore, the action-value function shares the parameter vector of
the reward function, with an associated feature vector $\mu^\pi$
called the \emph{feature expectation}. This notion will be of
primary importance for the contribution of this paper. Notice that
each component $\mu_i^\pi$ of this feature vector is actually the
action-value function of the policy $\pi$ assuming the reward is
$\phi_i$: $\mu_i^\pi(s,a) = Q^\pi_{\phi_i}(s,a)$. Therefore, any
algorithm designed for estimating an action-value function may be
used to estimate the feature expectation, such as Monte-Carlo
rollouts or temporal difference learning~\cite{Klein:2011}.

@@

Here, we consider a more restrictive set of classification
algorithms. We assume that the decision rule associates to an input
the argument which maximizes a related score function, this score
function being linearly parameterized and the associated parameters
being learnt by the algorithm. More formally, let $\psi(s,a) =
(\psi_1(x,y)  \dots  \psi_d(x,y))^\top\in \mathbb{R}^d$
%\begin{pmatrix}   \psi_1(x,y) & \dots & \psi_d(x,y)
%\end{pmatrix}^\top \in \mathbb{R}^d$
be a feature vector whose components are $d$ basis functions
$\psi_i\in\mathbb{R}^{\X\times\Y}$. The linearly parameterized score
function $s_w\in\mathbb{R}^{\X\times \Y}$ of parameter vector
$w\in\mathbb{R}^d$ is defined as $s_w(x,y) = w^\top \psi(x,y)$. The
associated decision rule $g_w\in{\Y^\X}$ is defined as $g_w(x) \in
\argmax_{y\in\Y}s_w(x,y)$. Using a training set $\{(x_i,y_i)_{1\leq
i\leq N}\}$, a linearly parameterized score function-based
multi-class classification (MC$^2$ for short) algorithm computes a
parameter vector $\theta_c$. The quality of the solution is
quantified by the classification error $\epsilon_c =
\prob(g_{\theta_c}(x)\neq y)$.

@@

Consider the classification framework of
Sec.~\ref{subsec:background:classif}. The input $x$ may be seen as a
state and the label $y$ as an action. Then, the decision rule
$g_w(x)$ can be interpreted as a policy which is greedy according to
the score function $w^\top \psi(x,y)$, which may itself be seen as
an action-value function. Making the parallel with
Eq.~\eqref{eq:def:mu}, if $\psi(x,y)$ is the feature expectation of
some policy $\pi$ which produces labels of the training set, and if
the classification error is small, then $w$ will be the parameter
vector of a reward function for which we may hope the policy $\pi$
to be near optimal. Based on these remarks, we're ready to present
the proposed Structured Classification-based IRL (SCIRL) algorithm.

Let $\pi_E$ be the expert policy from which we would like to recover
a reward function. Assume that we have a training set $\D =
\{(s_i,a_i=\pi_E(s_i))_{1\leq i\leq N}\}$ where states are sampled
according to the expert stationary distribution\footnote{For
example, if the Markov chain induced by the expert policy is
fast-mixing, sampling a trajectory will quickly lead to sample
states according to this distribution.} $\rho_E = \rho_{\pi_E}$.
Assume also that we have an estimate $\hat{\mu}^{\pi_E}$ of the
expert feature expectation $\mu^{\pi_E}$ defined in
Eq.~\eqref{eq:def:mu}. How to practically estimate this quantity is
postponed to Sec.~\ref{subsec:practicalApproach:muE}; however,
recall that estimating $\mu^{\pi_E}$ is simply a \emph{policy
evaluation} problem (estimating the action-value function of a given
policy), as noted in Sec.~\ref{subsec:background:irl}. Assume also
that an MC$^2$ algorithm has been chosen. The proposed algorithm
simply consists in choosing $\theta^\top\hat{\mu}^{\pi_E}(s,a)$ as
the linearly parameterized score function, training the classifier
on $\D$ which produces a parameter vector $\theta_c$, and outputting
the reward function $\R_{\theta_c}(s) = \theta_c^\top \phi(s)$.

\begin{algorithm2e}%[tbh]
    %\small
  \SetAlgoVlined
  \caption{SCIRL algorithm}
  \label{algo:scirl}
  %
  \BlankLine
  \emph{\textbf{Given}} a training set $\D = \{(s_i,a_i=\pi_E(s_i))_{1\leq i\leq N}\}$,
  an estimate $\hat{\mu}^{\pi_E}$ of the expert feature expectation $\mu^{\pi_E}$ and an MC$^2$ algorithm\;
  %
  \BlankLine
  \emph{\textbf{Compute}} the parameter vector $\theta_c$ using the
  MC$^2$ algorithm
  fed with the training set $\D$ and considering the parameterized score function
  $\theta^\top\hat{\mu}^{\pi_E}(s,a)$\;
  %
  \BlankLine
  \emph{\textbf{Output}} the reward function $\R_{\theta_c}(s) = \theta_c^\top\phi(s)$ \;
\end{algorithm2e}

The proposed approach is summarized in Alg.~\ref{algo:scirl}. We
call this Structured Classification-based IRL because using the
(estimated) expert feature expectation as the feature vector for the
classifier somehow implies taking into account the MDP structure
into the classification problem and allows outputting a reward
vector. Notice that contrary to most of existing IRL
%and apprenticeship learning
algorithms, SCIRL does not require solving the direct problem. If it
possibly requires estimating the expert feature expectation, it is
just a policy evaluation problem, less difficult than the policy
optimization issue involved by the direct problem. This is further
discussed in Sec.~\ref{sec:relatedWorks}.
%
%Now, we formally show that the proposed approach makes sens, as long
%as $\mu^{\pi_E}$ is well estimated and as long as the classification
%error is small.

@@

In this section, we show that the expert policy $\pi_E$ is close to
be optimal according to the reward function $\R_{\theta_c}$, more
precisely that
$\E_{s\sim\rho_E}[v^*_{\theta_c}(s)-v^{\pi_E}_{\theta_c}(s)]$ is
small. Before stating our main result, we need to introduce some
notations and to define some objects.

@@

\begin{theorem}
  \label{th}
  Let $\R_{\theta_c}$ be the reward function outputted by
  Alg.~\ref{algo:scirl}. Let also the quantities $C_f$, $\epsilon_c$
  and $\bar{\epsilon}_Q$ be defined as above. We have
  %
%  the concentration coefficient $C_f$,
%  the classification error $\epsilon_c$ and the mean-max
%  action-value function error $\bar{\epsilon}_Q$ be defined as
%  above. Then, we have
  \begin{equation}
    0\leq
    \E_{s\sim\rho_E}[v^*_{\R_{\theta_c}}-v^{\pi_E}_{\R_{\theta_c}}]
    \leq \frac{C_f}{1-\gamma}\left(\bar{\epsilon}_Q +
    \epsilon_c\frac{2\gamma\|\R_{\theta_c}\|_\infty}{1-\gamma}
%    \epsilon_c\frac{2\gamma}{1-\gamma}\min(\|\R_{\theta_c}\|_\infty,
%    C_p \E_{s\sim\rho_E}[\R_{\theta_c}(s)])
    \right).
  \end{equation}
\end{theorem}
\begin{proof}
  As the proof only relies on the reward $\R_{\theta_c}$, we
  omit the related subscripts to keep the notations simple
  (\textit{e.g.}, $v^\pi$ for
  $v^\pi_{\theta_c}=v^\pi_{\R_{\theta_c}}$ or $\R$ for $\R_{\theta_c}$). First, we link
  the error $\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)]$ to the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s)-v^{\pi_E}(s)]$.
  %
  % A second step consists in bounding this residual.
  %
  Componentwise, we have that:
  \begin{align}
    v^* - v^{\pi_E} &= T^* v^*  - T^{\pi^*}v^{\pi_E} +
    T^{\pi^*}v^{\pi_E} - T^* v^{\pi_E} + T^* v^{\pi_E} - v^{\pi_E}
    \\
    &\stackrel{(a)}{\leq} \gamma P_{\pi^*}(v^*-v^{\pi_E}) + T^*
    v^{\pi_E} - v^{\pi_E}
    %
    \stackrel{(b)}{\leq} (I-\gamma
    P_{\pi^*})^{-1} (T^* v^{\pi_E} - v^{\pi_E}).
  \end{align}
  Inequality $(a)$ holds because $T^{\pi^*} v^{\pi_E}\leq T^*
  v^{\pi_E}$ and inequality $(b)$ holds thanks
  to~\cite[Lemma~4.2]{Munos:2007}. Moreover, $v^*$ being optimal we
  have that $v^*-v^{\pi_E}\geq 0$ and $T^*$ being the Bellman
  optimality operator, we have $T^* v^{\pi_E}\geq
  T^{\pi_E}v^{\pi_E}=v^{\pi_E}$. Additionally, remark that
  $(I-\gamma P_{\pi^*})^{-1} = \sum_{t\geq 0}\gamma^t P_{\pi^*}^t$.
  Therefore, using the definition of the concentration coefficient
  $C_f$, we have that:
  \begin{equation}
    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq \frac{C_f}{1-\gamma}
    \E_{s\sim\rho_E}\left[[T^*v^{\pi_E}](s) - v^{\pi_E}(s)\right].
    \label{eq:proof:residual}
  \end{equation}
  This results actually follows closely the one
  of~\cite[Theorem~4.2]{Munos:2007}. There remains to bound the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s) -
  v^{\pi_E}(s)]$. Considering the following decomposition,
  \begin{equation}
    T^* v^{\pi_E} - v^{\pi_E} = T^* v^{\pi_E} - T^{\pi_c}v^{\pi_E}
    + T^{\pi_c}v^{\pi_E}- v^{\pi_E},
    \label{eq:proof:decomposition}
  \end{equation}
  we will bound $\E_{s\sim\rho_E}[[T^* v^{\pi_E}](s) - [T^{\pi_c}v^{\pi_E}](s)]$
  and $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]$.

  The policy $\pi_c$ (the decision rule of the classifier) is greedy
  with respect to
  $\hat{Q}^{\pi_E}=\theta_c^\top\hat{\mu}^{\pi_E}$. Therefore, for any state-action couple
  $(s,a)\in\s\times \A$ we have:
  \begin{equation}
    \hat{Q}^{\pi_E}(s,\pi_c(s))\geq
    \hat{Q}^{\pi_E}(s,a)
    \Leftrightarrow
    Q^{\pi_E}(s,a) \leq Q^{\pi_E}(s,\pi_c(s)) +
    \epsilon_Q(s,\pi_c(s)) - \epsilon_Q(s,a).
  \end{equation}
  By definition, $Q^{\pi_E}(s,a) = [T^a v^{\pi_E}](s)$ and
  $Q^{\pi_E}(s,\pi_c(s)) = [T^{\pi_c} v^{\pi_E}](s)$. Therefore, for $s\in\s$:
  \begin{align}
    \forall a\in A,\; [T^a v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \epsilon_Q(s,\pi_c(s))-\epsilon_Q(s,a)
    \\
    \Rightarrow [T^* v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \max_{a\in \A}\epsilon_Q(s,a)-\min_{a\in
    \A}\epsilon_Q(s,a).
  \end{align}
  Taking the expectation according to $\rho_E$ and noticing that
  $T^* v^{\pi_E}\geq v^{\pi_E}$, we bound the first
  term:
  \begin{equation}
    0 \leq \E_{s\sim\rho_E}\left[ [T^* v^{\pi_E}](s) - [T^{\pi_c}
    v^{\pi_E}](s)\right] \leq \bar{\epsilon}_Q.
    \label{eq:proof:b1}
  \end{equation}
  There finally remains to bound the term $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) -
  v^{\pi_E}(s)]$.

>>>>

\begin{document}
\newtheorem{prop}[theorem]{Proposition}
\maketitle

@@

\section{Contexte et notations}
\label{sec:background}

\subsection{Apprentissage par Renforcement (Inverse)}
\label{subsec:background:irl}

Un Processus Décisionnel de Markov (PDM)~\cite{Puterman:1994} est un ensemble $\{\s,\A,\p,\R,\gamma\}$ avec $\s$ l'espace d'état fini\footnote{Cette contribution peut être étendue aux espaces d'état compacts moyennant le traitement de certains détails techniques.}, $\A$ l'espace d'action fini, $\p =
\{P_a = (p(s'|s,a))_{1\leq s,s'\leq |\s|}, a\in\A\}$ l'ensemble des probabilités de transition markoviennes, $\R\in\mathbb{R}^\s$ la fonction de récompense sur les états et $\gamma$ le facteur d'oubli.
Une politique déterministe $\pi\in\s^\A$ définit le comportement d'un agent. On quantifie la qualité de ce comportement par le biais de la fonction de valeur $v_\R^\pi\in\mathbb{R}^\s$ qui à chaque état associe la récompense pondérée cumulée recueillie par l'agent lorsqu'il part de cet état et suit la politique $\pi$ par la suite : $v_\R^\pi(s) = \E[\sum_{t\geq 0} \gamma^t \R(S_t)|S_0=s,\pi]$. Une politique optimale $\pi_\R^*$ (vis-à-vis de la fonction de récompense $\R$) est une politique dont la fonction de valeur $v^*_\R$ satisfait $v_\R^* \geq v_\R^\pi$, composante par composante, pour toute politique $\pi$.

Soit $P_\pi$ la matrice stochastique $P_\pi =
(p(s'|s,\pi(s)))_{1\leq s,s'\leq |\s|}$. Un léger abus de notation nous permet de nommer $a$ la politique associant l'action $a$ à chaque état $s$. Les opérateurs d'évaluation (resp. d'optimalité) de Bellman $T^\pi_\R\text{ (resp. $T^*_\R$)}:\mathbb{R}^\s
\rightarrow \mathbb{R}^\s$ sont définis par $T^\pi_\R v = \R + \gamma
P_\pi v$ and $T_\R^*v = \max_\pi T_\R^\pi v$.
%\begin{equation}
%  T^\pi_\R v = \R + \gamma P_\pi v \text{ and } T_\R^*v = \max_\pi T_\R^\pi v.
%\end{equation}
Ces opérateurs sont des contractions qui admettent $v_\R^\pi$ et $v^*_\R$ comme points fixes respectifs : $v_\R^\pi = T^\pi_\R v_\R^\pi$ et $v^*_\R = T^*_\R v^*_\R$.
La fonction de qualité $Q^\pi\in\mathbb{R}^{\s\times \A}$ ajoute à $v$ un degré de liberté dans le choix de la première action, on la définit formellement par $Q_\R^\pi(s,a)
= [T^a_\R v^\pi_\R](s)$. On note $\rho_\pi$ la distribution stationnaire de la politique $\pi$ (qui satisfait $\rho_\pi^\top P_\pi = \rho_\pi^\top$).

L'Apprentissage par Renforcement (AR) et la programmation dynamique approchée cherchent à estimer la politique de contrôle optimale lorsque le modèle (les probabilités de transition et la fonction de récompense) est inconnu (mais observé par interaction avec le système à contrôler) et quand l'espace d'état est trop grand pour qu'une représentation exacte des objets concernés (fonctions de valeur, politiques) soit possible~\cite{Bertsekas:1996,Sutton:1998,szepesvari2010c}.
%
On appelle cela le problème direct.
%
L'Apprentissage par Renforcement Inverse (approché)~\cite{Ng:2000} cherche à estimer la fonction de récompense pour laquelle une politique observée est optimale.
%
On nomme cette politique la politique de l'expert, notée $\pi_E$. On suppose qu'elle est optimale vis-à-vis d'une certaine fonction de récompense $R_E$ inconnue. Le but de l'ARI est de calculer une fonction de récompense $\hat{\R}$ telle que la politique de l'expert soit (quasiment) optimale, c'est-à-dire telle que $v^*_{\hat{\R}}
\approx v^{\pi_E}_{\hat{\R}}$.
%
On appelle cela le problème inverse.

De la même manière que pour le problème direct, l'espace d'état peut être trop grand pour que la fonction de récompense admette une représentation exacte exploitable. On se limite donc de fait à la recherche d'une bonne fonction de récompense paramétrée linéairement. Soit $\phi(s) = (\phi_1(s)  \dots
\phi_p(s))^\top$
%\begin{pmatrix}
%  \phi_1(s) & \dots & \phi_p(s)
%\end{pmatrix}^\top$
un vecteur d'attributs composé de $p$ fonctions de base $\phi_i\in\mathbb{R}^\s$, on définit les fonctions de récompense paramétrées par $\R_\theta(s) = \theta^\top \phi(s) = \sum_{i=1}^p
\theta_i \phi_i(s)$.
%\begin{equation}
%  \R_\theta(s) = \theta^\top \phi(s) = \sum_{i=1}^p \theta_i \phi_i(s).
%\end{equation}
La recherche d'une bonne récompense est donc ramenée à la recherche d'un bon vecteur de paramètres vector $\theta \in\mathbb{R}^p$. Nous utiliserons indifféremment $\R_\theta$ et $\theta$ comme indices (\textit{e.g.}, $v_\theta^\pi$ pour $v_{\R_\theta}^\pi$).
Cette paramétrisation de la récompense implique une paramétrisation similaire de la fonction de qualité :
\begin{equation}
  Q^\pi_\theta(s,a) = \theta^\top \mu^\pi(s,a) \text{ with }
  \mu^\pi(s,a) = \E[\sum_{t\geq 0} \gamma^t
  \phi(S_t)|S_0=s,A_0=a,\pi].
  \label{eq:def:mu}
\end{equation}
Conséquemment, la fonction de qualité partage son vecteur de paramètres avec la fonction de récompense, mais en relation avec un vecteur d'attributs $\mu^\pi$ appelé l'attribut moyen. Cette notion est de prime importance à notre contribution. Chaque composante $\mu_i^\pi$ de ce vecteur d'attributs est en réalité une fonction de qualité de la politique $\pi$ où la récompense serait $\phi_i$: $\mu_i^\pi(s,a) = Q^\pi_{\phi_i}(s,a)$. De fait, tout algorithme d'estimation de la fonction de qualité peut être utilisé pour estimer l'attribut moyen, comme une estimation de Monte-Carlo ou un algorithme aux différences temporelles~\cite{Klein:2011}.

@@

Here, we consider a more restrictive set of classification
algorithms. We assume that the decision rule associates to an input
the argument which maximizes a related score function, this score
function being linearly parameterized et the associated parameters
being learnt by the algorithm. More formally, let $\psi(s,a) =
(\psi_1(x,y)  \dots  \psi_d(x,y))^\top\in \mathbb{R}^d$
%\begin{pmatrix}   \psi_1(x,y) & \dots & \psi_d(x,y)
%\end{pmatrix}^\top \in \mathbb{R}^d$
be a feature vector whose components are $d$ basis functions
$\psi_i\in\mathbb{R}^{\X\times\Y}$. The linearly parameterized score
function $s_w\in\mathbb{R}^{\X\times \Y}$ of parameter vector
$w\in\mathbb{R}^d$ is defined as $s_w(x,y) = w^\top \psi(x,y)$. The
associated decision rule $g_w\in{\Y^\X}$ is defined as $g_w(x) \in
\argmax_{y\in\Y}s_w(x,y)$. Using a training set $\{(x_i,y_i)_{1\leq
i\leq N}\}$, a linearly parameterized score function-based
multi-class classification (MC$^2$ for short) algorithm computes a
parameter vector $\theta_c$. The quality of the solution is
quantified by the classification error $\epsilon_c =
\prob(g_{\theta_c}(x)\neq y)$.

@@

Consider the classification framework of
Sec.~\ref{subsec:background:classif}. The input $x$ may be seen as a
state et the label $y$ as an action. Then, the decision rule
$g_w(x)$ can be interpreted as a policy which is greedy according to
the score function $w^\top \psi(x,y)$, which may itself be seen as
an action-value function. Making the parallel with
Eq.~\eqref{eq:def:mu}, if $\psi(x,y)$ is the feature expectation of
some policy $\pi$ which produces labels of the training set, et if
the classification error is small, then $w$ will be the parameter
vector of a reward function for which we may hope the policy $\pi$
to be near optimal. Based on these remarks, we're ready to present
the proposed Structured Classification-based IRL (SCIRL) algorithm.

Let $\pi_E$ be the expert policy from which we would like to recover
a reward function. Assume that we have a training set $\D =
\{(s_i,a_i=\pi_E(s_i))_{1\leq i\leq N}\}$ where states are sampled
according to the expert stationary distribution\footnote{For
example, if the Markov chain induced by the expert policy is
fast-mixing, sampling a trajectory will quickly lead to sample
states according to this distribution.} $\rho_E = \rho_{\pi_E}$.
Assume also that we have an estimate $\hat{\mu}^{\pi_E}$ of the
expert feature expectation $\mu^{\pi_E}$ defined in
Eq.~\eqref{eq:def:mu}. How to practically estimate this quantity is
postponed to Sec.~\ref{subsec:practicalApproach:muE}; however,
recall that estimating $\mu^{\pi_E}$ is simply a \emph{policy
evaluation} problem (estimating the action-value function of a given
policy), as noted in Sec.~\ref{subsec:background:irl}. Assume also
that an MC$^2$ algorithm has been chosen. The proposed algorithm
simply consists in choosing $\theta^\top\hat{\mu}^{\pi_E}(s,a)$ as
the linearly parameterized score function, training the classifier
on $\D$ which produces a parameter vector $\theta_c$, et outputting
the reward function $\R_{\theta_c}(s) = \theta_c^\top \phi(s)$.

\begin{algorithm2e}%[tbh]
    %\small
  \SetAlgoVlined
  \caption{SCIRL algorithm}
  \label{algo:scirl}
  %
  \BlankLine
  \emph{\textbf{Given}} a training set $\D = \{(s_i,a_i=\pi_E(s_i))_{1\leq i\leq N}\}$,
  an estimate $\hat{\mu}^{\pi_E}$ of the expert feature expectation $\mu^{\pi_E}$ et an MC$^2$ algorithm\;
  %
  \BlankLine
  \emph{\textbf{Compute}} the parameter vector $\theta_c$ using the
  MC$^2$ algorithm
  fed with the training set $\D$ et considering the parameterized score function
  $\theta^\top\hat{\mu}^{\pi_E}(s,a)$\;
  %
  \BlankLine
  \emph{\textbf{Output}} the reward function $\R_{\theta_c}(s) = \theta_c^\top\phi(s)$ \;
\end{algorithm2e}

The proposed approach is summarized in Alg.~\ref{algo:scirl}. We
call this Structured Classification-based IRL because using the
(estimated) expert feature expectation as the feature vector for the
classifier somehow implies taking into account the MDP structure
into the classification problem et allows outputting a reward
vector. Notice that contrary to most of existing IRL
%and apprenticeship learning
algorithms, SCIRL does not require solving the direct problem. If it
possibly requires estimating the expert feature expectation, it is
just a policy evaluation problem, less difficult than the policy
optimization issue involved by the direct problem. This is further
discussed in Sec.~\ref{sec:relatedWorks}.
%
%Now, we formally show that the proposed approach makes sens, as long
%as $\mu^{\pi_E}$ is well estimated et as long as the classification
%error is small.

@@

In this section, we show that the expert policy $\pi_E$ is close to
be optimal according to the reward function $\R_{\theta_c}$, more
precisely that
$\E_{s\sim\rho_E}[v^*_{\theta_c}(s)-v^{\pi_E}_{\theta_c}(s)]$ is
small. Before stating our main result, we need to introduce some
notations et to define some objects.

@@

\begin{theorem}
  \label{th}
  Let $\R_{\theta_c}$ be the reward function outputted by
  Alg.~\ref{algo:scirl}. Let also the quantities $C_f$, $\epsilon_c$
  et $\bar{\epsilon}_Q$ be defined as above. We have
  %
%  the concentration coefficient $C_f$,
%  the classification error $\epsilon_c$ et the mean-max
%  action-value function error $\bar{\epsilon}_Q$ be defined as
%  above. Then, we have
  \begin{equation}
    0\leq
    \E_{s\sim\rho_E}[v^*_{\R_{\theta_c}}-v^{\pi_E}_{\R_{\theta_c}}]
    \leq \frac{C_f}{1-\gamma}\left(\bar{\epsilon}_Q +
    \epsilon_c\frac{2\gamma\|\R_{\theta_c}\|_\infty}{1-\gamma}
%    \epsilon_c\frac{2\gamma}{1-\gamma}\min(\|\R_{\theta_c}\|_\infty,
%    C_p \E_{s\sim\rho_E}[\R_{\theta_c}(s)])
    \right).
  \end{equation}
\end{theorem}
\begin{proof}
  As the proof only relies on the reward $\R_{\theta_c}$, we
  omit the related subscripts to keep the notations simple
  (\textit{e.g.}, $v^\pi$ for
  $v^\pi_{\theta_c}=v^\pi_{\R_{\theta_c}}$ or $\R$ for $\R_{\theta_c}$). First, we link
  the error $\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)]$ to the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s)-v^{\pi_E}(s)]$.
  %
  % A second step consists in bounding this residual.
  %
  Componentwise, we have that:
  \begin{align}
    v^* - v^{\pi_E} &= T^* v^*  - T^{\pi^*}v^{\pi_E} +
    T^{\pi^*}v^{\pi_E} - T^* v^{\pi_E} + T^* v^{\pi_E} - v^{\pi_E}
    \\
    &\stackrel{(a)}{\leq} \gamma P_{\pi^*}(v^*-v^{\pi_E}) + T^*
    v^{\pi_E} - v^{\pi_E}
    %
    \stackrel{(b)}{\leq} (I-\gamma
    P_{\pi^*})^{-1} (T^* v^{\pi_E} - v^{\pi_E}).
  \end{align}
  Inequality $(a)$ holds because $T^{\pi^*} v^{\pi_E}\leq T^*
  v^{\pi_E}$ et inequality $(b)$ holds thanks
  to~\cite[Lemma~4.2]{Munos:2007}. Moreover, $v^*$ being optimal we
  have that $v^*-v^{\pi_E}\geq 0$ et $T^*$ being the Bellman
  optimality operator, we have $T^* v^{\pi_E}\geq
  T^{\pi_E}v^{\pi_E}=v^{\pi_E}$. Additionally, remark that
  $(I-\gamma P_{\pi^*})^{-1} = \sum_{t\geq 0}\gamma^t P_{\pi^*}^t$.
  Therefore, using the definition of the concentration coefficient
  $C_f$, we have that:
  \begin{equation}
    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq \frac{C_f}{1-\gamma}
    \E_{s\sim\rho_E}\left[[T^*v^{\pi_E}](s) - v^{\pi_E}(s)\right].
    \label{eq:proof:residual}
  \end{equation}
  This results actually follows closely the one
  of~\cite[Theorem~4.2]{Munos:2007}. There remains to bound the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s) -
  v^{\pi_E}(s)]$. Considering the following decomposition,
  \begin{equation}
    T^* v^{\pi_E} - v^{\pi_E} = T^* v^{\pi_E} - T^{\pi_c}v^{\pi_E}
    + T^{\pi_c}v^{\pi_E}- v^{\pi_E},
    \label{eq:proof:decomposition}
  \end{equation}
  we will bound $\E_{s\sim\rho_E}[[T^* v^{\pi_E}](s) - [T^{\pi_c}v^{\pi_E}](s)]$
  et $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]$.

  The policy $\pi_c$ (the decision rule of the classifier) is greedy
  with respect to
  $\hat{Q}^{\pi_E}=\theta_c^\top\hat{\mu}^{\pi_E}$. Therefore, for any state-action couple
  $(s,a)\in\s\times \A$ we have:
  \begin{equation}
    \hat{Q}^{\pi_E}(s,\pi_c(s))\geq
    \hat{Q}^{\pi_E}(s,a)
    \Leftrightarrow
    Q^{\pi_E}(s,a) \leq Q^{\pi_E}(s,\pi_c(s)) +
    \epsilon_Q(s,\pi_c(s)) - \epsilon_Q(s,a).
  \end{equation}
  By definition, $Q^{\pi_E}(s,a) = [T^a v^{\pi_E}](s)$ and
  $Q^{\pi_E}(s,\pi_c(s)) = [T^{\pi_c} v^{\pi_E}](s)$. Therefore, for $s\in\s$:
  \begin{align}
    \forall a\in A,\; [T^a v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \epsilon_Q(s,\pi_c(s))-\epsilon_Q(s,a)
    \\
    \Rightarrow [T^* v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \max_{a\in \A}\epsilon_Q(s,a)-\min_{a\in
    \A}\epsilon_Q(s,a).
  \end{align}
  Taking the expectation according to $\rho_E$ et noticing that
  $T^* v^{\pi_E}\geq v^{\pi_E}$, we bound the first
  term:
  \begin{equation}
    0 \leq \E_{s\sim\rho_E}\left[ [T^* v^{\pi_E}](s) - [T^{\pi_c}
    v^{\pi_E}](s)\right] \leq \bar{\epsilon}_Q.
    \label{eq:proof:b1}
  \end{equation}
  There finally remains to bound the term $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) -
  v^{\pi_E}(s)]$.
